Lucky Thirteen - A Cryptographic Analyses

### 2013-07-14 paper.tex
Section V and IV-B alpha, Sections I, II, III, IV-A beta

% SECTION II-B
\subsection{MAC - Message Authentication Code}
\label{sec:crypto:hmac}
To ensure message integrity, RFC5246~\cite{ietf2008transport} prescribes the use of a keyed Message Authentication Code (MAC). The MAC cipher suites described in the RFC are based on hashing functions, HMACs, and are described in~\cite{krawczyk1997rfc}. Though, the RFC does not prohibit the use of other cipher suites, the implementations found by~\citeauthor{alfardan2013lucky} in~\cite{alfardan2013lucky} all restrict the use to HMAC-MD5, HMAC-SHA-1 or HMAC-SHA-256~\footnote{HMAC-MD5 produces 16 bytes of MAC.\\HMAC-SHA-1 produces 20 bytes of MAC.\\HMAC-SHA-256 produces 32 bytes of MAC.} To compute the HMAC for a message M and key K, the specified HMAC algorithms use the following formula. 
\[ \text{HMAC}(K,M) = H((K \oplus opad)||H(K \oplus ipad)||M)). \]
With \textit{opad}, \textit{ipad} to be specific $64$-byte values and key $K$ zero-padded to $64$ bytes. "Each hash function uses a \textit{Merkle-Damg\aa rd strengthening}. Here, an 8-byte length field followed by padding of a specified byte format are appended to the message M to be hashed. The padding is at least $1$ byte in length and aligns the data on a $64$-byte boundary. The relevant hash functions also have an iterated structure, processing messages in chunks of $64$ bytes ($512$ bits) using a compression function, with the output of each compression step being chained into the next step. The compression function in turn involves a complex round structure, with many basic arithmetic operations on data being involved in each round."~\cite{alfardan2013lucky}.\footnote{The exact working of this Merkle-Damg\aa rd strengthening is not important for the Lucky Thirteen and therefore falls out of the scope of this paper.}
This means that, depending on the size of message M, the number of runs of the hash function H can be computed. Because of the Merkle-Damg\aa rd strengthening, a message M is appended by the $8$-byte length field, and padded to be a multiple of $64$ bytes. This means that the first of the Merkle-Damg\aa rd rounds consists of $55$ bytes of message M, followed by $64$ bytes for every following $64$ bytes of message M. The inner hash function depends on the $\text{message length of} M + 1$ constant hash function round, the outer hash function is a constant of $2$. This means that the following formula represents the number of hash function calls for L is the message length of M in bytes:
\[ \lceil \frac{L - 55}{64} \rceil + 4 \]

@@

% SECTION IV-B
\subsection{Bomb Oracles}
\label{sec:paddingoracle:bomb}
Problem: \texttt{decryption\_failed} is a fatal error. \\
\textcolor{red}{TODO: fixed using D, with multi-session TLS attacks}

>>>>

% SECTION II-B
\subsection{MAC - Message Authentication Code}
\label{sec:crypto:hmac}
To ensure message integrity, RFC5246~\cite{ietf2008transport} prescribes the use of a keyed Message Authentication Code (MAC). The MAC cipher suites described in the RFC are based on hashing functions, HMACs, and are described in~\cite{krawczyk1997rfc}. Though, the RFC does not prohibit the use of other cipher suites, the implementations found by~\citeauthor{alfardan2013lucky} in~\cite{alfardan2013lucky} all restrict the use to HMAC-MD5, HMAC-SHA-1 or HMAC-SHA-256~\footnote{HMAC-MD5 produces 16 bytes of MAC.\\HMAC-SHA-1 produces 20 bytes of MAC.\\HMAC-SHA-256 produces 32 bytes of MAC.} To compute the HMAC for a message M and key K, the specified HMAC algorithms use the following formula. 
\[ \text{HMAC}(K,M) = H((K \oplus opad)||H(K \oplus ipad)||M)). \]
With \textit{opad}, \textit{ipad} to be specific $64$-byte values and key $K$ zero-padded to $64$ bytes. "Each hash function uses a \textit{Merkle-Damg\aa rd strengthening}. Here, an 8-byte length field followed by padding of a specified byte format are appended to the message M to be hashed. The padding is at least $1$ byte in length and aligns the data on a $64$-byte boundary. The relevant hash functions also have an iterated structure, processing messages in chunks of $64$ bytes ($512$ bits) using a compression function, with the output of each compression step being chained into the next step. The compression function in turn involves a complex round structure, with many basic arithmetic operations on data being involved in each round."~\cite{alfardan2013lucky}.\footnote{The exact working of this Merkle-Damg\aa rd strengthening is not important for this paper and therefore falls out of the scope of this paper.}
This means that, depending on the size of message M, the number of runs of the hash function H can be computed. Because of the Merkle-Damg\aa rd strengthening, a message M is appended by the $8$-byte length field, and padded to be a multiple of $64$ bytes. This means that the first of the Merkle-Damg\aa rd rounds consists of $55$ bytes of message M, followed by $64$ bytes for every following $64$ bytes of message M. The inner hash function depends on the $\text{message length of} M + 1$ constant hash function round, the outer hash function is a constant of $2$. This means that the following formula represents the number of hash function calls for L is the message length of M in bytes:
\[ \lceil \frac{L - 55}{64} \rceil + 4 \]

@@

% SECTION IV-B
\subsection{Bomb Oracles}
\label{sec:paddingoracle:bomb}
As discussed in Section~\ref{sec:paddingoracle:padding}, the attack by~\citeauthor{vaudenay2002security}~\cite{vaudenay2002security} would only work if the TLS connections would not have been terminated at the first wrongly padded plaintext received by the Oracle. The authors of a novel attack that includes multi-session byte querying~\cite{canvel2003password} have solved this issue by the introduction of so called \textit{bomb oracles}, which explode at receiving a wrongly padded plaintext. Their attack makes use of two following requirements.
\begin{itemize}
	\item The attacker is able to re-initiate the TLS session after it terminates.
	\item The secret plaintext $P^{*}$ appears in the same position in each additional TLS connection.
\end{itemize}
Fortunately, these two requirements can be met by a multi-session attack~\cite{vaudenay2002security}. The attack works by feeding the browser of some user with specially crafted Javascript that repeatedly connects to the Oracle using a TLS connection. With every new connection that is set-up, the cookie ($P^{*}$) is at a known location in the HTTPS stream, see Figure~\ref{fig:beast_http}\footnote{For further information about the HTTPS stream, the reader is referred to the paper explaining the BEAST attack~\cite{duong2011here}, it is considered out of scope for this paper}. In this figure, the authors of~\cite{duong2011here} have an exemplary block size of 8 bytes. They show that by controlling some of the bytes within the HTTPS stream, the bytes that the attacker is after can be forced towards a fixed position at the starting point of a new block.  This way of attacking is called a \textit{blockwise chosen-boundary attack (BCBA)}~\cite{duong2011here}.

\begin{figure}[h]
	\centering
	\includegraphics[width=0.5\textwidth]{beast_https.png}
	\caption{HTTPS Record.~\cite{duong2011here}}
	\label{fig:beast_http}
\end{figure}

While having plaintext block $P^{*}$ at the same, fixed position at the beginning of a new block, an attacker can use the following novelty to extract the plaintext while it is encrypted with different keys by querying for assumed plaintext bytes.

Suppose that the attacker has the possibility of finding $C^{*}$ and $C^{'}$ (preceding $C^{*}$) in a TLS connection. The attacker can guess what the plaintext $P^{*}$ corresponding to $C^{*}$ is, and test this hypothesis by recalling the following. (This should be done for all possible plaintexts).
\[
\begin{split}
C^{*} &= E_K(P^{*} \oplus C^{'}) \\
P^{*} &= D_K(C^{*}) \oplus C^{'}
\end{split}
\]
If an attacker wants to manipulate a byte to become some part of the padding (of importance as explained thoroughly in Section~\ref{sec:paddingoracle:padding}), the attacker can craft the wanted byte as follows. Consider the attacker wants to turn the i'th byte in the decrypted $P$ into 0x(15-i), if he guesses the i'th byte of $P^{*}$ is $P_i^{*}$.
\[ D_K(C_i^{*}) = P_i^{*} \oplus 0\text{x}(15-i) \oplus C_i^{'} \]
Thus he can use the same technique as used in Section~\ref{sec:paddingoracle:padding} with $\Delta_i = P_i^{*} \oplus 0\text{x}(15-i)$, and set al $\Delta_j$'s $j>i$ to also match the correct padding. With every new TLS connection set-up, the attacker needs to receive the $C^{'}$ and $C^{*}$ again and build the query for the Oracle with a new pair of these, though, the two ciphertexts he needs for this are sent in the set-up and thus can be easily extracted.

The advantage of using this technique will be elaborated upon in Section~\ref{sec:lucky:improvements}, but one can image that the different guesses for $P_i^{*}$ can be ordered for probability, instead of checking all plaintexts, speeding up this attack considerably.
